---
title: "ardlverse: An R Ecosystem for ARDL Modeling and Cointegration Analysis"
abstract: >
  We introduce the ardlverse, an ecosystem of R packages for Autoregressive
  Distributed Lag (ARDL) modeling and bounds testing for cointegration. The
  ecosystem includes packages for augmented ARDL (aardl), rolling and recursive
  ARDL (rardl), Fourier bootstrap NARDL (fbnardl), multiple-threshold NARDL
  (mtnardl), Fourier quantile ARDL (fqardl), quantile ARDL (qardlr), panel
  NARDL (pnardl), panel quantile ARDL (xtpqardl), and cross-sectionally
  augmented panel quantile ARDL (xtcspqardl). These packages substantially
  extend the ARDL bounds testing framework of Pesaran, Shin, and Smith (2001)
  in several directions: bootstrap inference, nonlinear decompositions, Fourier
  approximations for smooth structural breaks, quantile regression extensions,
  and panel data variants with cross-sectional dependence correction. Together,
  they provide the most comprehensive ARDL toolkit available in any statistical
  software environment. All packages are open-source under the GPL-3 license.
author:
  - name: Muhammad Abdullah Alkhalaf
    affiliation: Rufyq Elngeh for Academic and Business Services
    address: Riyadh, Saudi Arabia
    orcid: 0009-0002-2677-9246
    email: muhammedalkhalaf@gmail.com
type: package
output:
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: paper.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = "#>")
```

# Introduction

The Autoregressive Distributed Lag (ARDL) bounds test of \citet{PSS2001} is among
the most widely used cointegration frameworks in applied economics. It permits
testing for long-run relationships regardless of whether regressors are I(0),
I(1), or mutually cointegrated. The existing R implementation in the `ARDL`
package \citep{Natsiopoulos2022} provides standard linear ARDL estimation and
bounds testing but lacks bootstrap bounds testing, nonlinear decompositions,
Fourier approximations, quantile extensions, and panel variants.

The `ardlverse` ecosystem fills these gaps with nine packages spanning the full
range of ARDL methodology. This article describes each package, demonstrates
usage with simulated data, and compares with existing implementations.

# The aardl package

The `aardl` package implements the augmented ARDL of \citet{Sam2019}, which
resolves degenerate cases in the PSS bounds test by adding F-overall, t-dependent,
and F-independent test statistics. It supports linear, bootstrap \citep{McNown2018},
Fourier, and NARDL variants.

```{r aardl-example}
library(ardlverse)

# Simulate cointegrated time series
set.seed(42)
n <- 200
x1 <- cumsum(rnorm(n))
x2 <- cumsum(rnorm(n))
e <- rnorm(n, 0, 0.5)
y <- 0.5 + 0.8 * x1 - 0.3 * x2 + e

ts_data <- data.frame(y = y, x1 = x1, x2 = x2)

# Augmented ARDL with bootstrap bounds test
result <- aardl(y ~ x1 + x2, data = ts_data, p = 4, q = 4,
                case = 3, type = "bootstrap", nboot = 2000)
summary(result)
```

# The rardl package

The `rardl` package implements rolling-window and recursive ARDL bounds testing
\citep{Shahbaz2023, Khan2023} for detecting structural instability in
cointegrating relationships over time.

```{r rardl-example}
# Rolling-window ARDL
result <- rardl(y ~ x1 + x2, data = ts_data,
                method = "rolling", window = 60, case = 3)
plot(result)
```

The `plot` method produces time-varying F-statistics with confidence bands,
allowing visual identification of periods where the cointegrating relationship
breaks down.

# The fbnardl package

The `fbnardl` package implements Fourier NARDL with \citet{Kripfganz2020}
critical values and Fourier Bootstrap NARDL with bootstrap cointegration tests
\citep{Bertelli2022}. Fourier terms capture smooth structural breaks without
requiring break date specification.

```{r fbnardl-example}
library(fbnardl)

set.seed(123)
n <- 250
x1 <- cumsum(rnorm(n))
y <- 0.5 * pmax(diff(c(0, x1)), 0) - 0.2 * pmin(diff(c(0, x1)), 0) +
     cumsum(rnorm(n, 0, 0.3))
ts_data <- data.frame(y = y, x1 = x1)

# Fourier Bootstrap NARDL
result <- fbnardl(y ~ x1, data = ts_data, decompose = "x1",
                  type = "fbnardl", maxlag = 4, maxk = 3, reps = 999)
summary(result)
```

# The mtnardl package

The `mtnardl` package implements the multiple-threshold NARDL of \citet{Pal2016},
which decomposes regressors into regime-specific partial sums based on multiple
thresholds rather than just the zero threshold of standard NARDL.

```{r mtnardl-example}
# Multiple-threshold NARDL
result <- mtnardl(y ~ x1, data = ts_data,
                  thresholds = c(-0.5, 0, 0.5), case = 3)
summary(result)
```

# The fqardl package

The `fqardl` package combines Fourier terms with quantile regression in the ARDL
framework, allowing both smooth structural change and distributional heterogeneity.

```{r fqardl-example}
library(fqardl)

set.seed(456)
n <- 200
x1 <- cumsum(rnorm(n))
y <- 0.3 * x1 + rnorm(n)
ts_data <- data.frame(y = y, x1 = x1)

result <- fqardl(y = ts_data$y, x = ts_data$x1,
                 tau = c(0.25, 0.5, 0.75), type = "fqardl")
summary(result)
```

# The qardlr package

The `qardlr` package implements the quantile ARDL of \citet{Cho2015} with
quantile-specific long-run, short-run, and impact parameters, BIC-based lag
selection, and Wald tests for parameter constancy across quantiles.

```{r qardlr-example}
library(qardlr)

set.seed(789)
n <- 200
x1 <- cumsum(rnorm(n))
x2 <- cumsum(rnorm(n))
y <- 0.5 * x1 - 0.3 * x2 + rnorm(n)
ts_data <- data.frame(y = y, x1 = x1, x2 = x2)

result <- qardl(y ~ x1 + x2, data = ts_data,
                tau = seq(0.1, 0.9, 0.1), p = 2, q = 2)
summary(result)
```

# The pnardl package

The `pnardl` package extends NARDL \citep{Shin2014} to panel data with Pooled
Mean Group (PMG), Mean Group (MG), and Dynamic Fixed Effects (DFE) estimators.

```{r pnardl-example}
# Simulate panel data
set.seed(101)
N <- 10; T_obs <- 50
panel_data <- data.frame(
  country = rep(paste0("C", 1:N), each = T_obs),
  year = rep(2001:(2000 + T_obs), N),
  y = unlist(lapply(1:N, function(i) {
    x <- cumsum(rnorm(T_obs))
    0.5 * pmax(diff(c(0, x)), 0) - 0.2 * pmin(diff(c(0, x)), 0) +
      rnorm(T_obs, 0, 0.5)
  })),
  x1 = unlist(lapply(1:N, function(i) cumsum(rnorm(T_obs))))
)

# Panel NARDL with PMG estimator
result <- pnardl(y ~ x1, data = panel_data,
                 id = "country", time = "year",
                 p = 1, q = 1)
summary(result)
```

# The xtpqardl package

The `xtpqardl` package implements panel quantile ARDL with PMG, MG, and DFE
estimators across multiple quantiles \citep{Cho2015, Bildirici2022}.

```{r xtpqardl-example}
library(xtpqardl)

result <- xtpqardl(y ~ x1, data = panel_data,
                   id = "country", time = "year", lr = c("x1"),
                   tau = c(0.25, 0.5, 0.75), model = "pmg")
summary(result)
```

# The xtcspqardl package

The `xtcspqardl` package implements cross-sectionally augmented panel quantile
ARDL with QCCEMG and CS-PQARDL estimators for panel data with cross-sectional
dependence \citep{Pesaran2006cce, Chudik2015}.

```{r xtcspqardl-example}
library(xtcspqardl)

result <- xtcspqardl(y ~ x1, data = panel_data,
                     id = "country", time = "year",
                     tau = 0.5, estimator = "qccemg")
summary(result)
```

# Comparison with existing implementations

The `ARDL` package \citep{Natsiopoulos2022} is the primary existing R
implementation for ARDL bounds testing. It provides standard linear ARDL
estimation, automatic lag selection, and the PSS F- and t-bounds tests. However,
it does not support:

- **Bootstrap bounds testing** (provided by `aardl` via \citealt{McNown2018})
- **Augmented ARDL** for degenerate cases (provided by `aardl` via \citealt{Sam2019})
- **Rolling/recursive ARDL** for structural stability (provided by `rardl`)
- **Nonlinear ARDL** with asymmetric decompositions (provided by `aardl`, `fbnardl`, `mtnardl`, `pnardl`)
- **Fourier approximations** for smooth breaks (provided by `fbnardl`, `fqardl`)
- **Quantile ARDL** (provided by `qardlr`, `fqardl`, `xtpqardl`, `xtcspqardl`)
- **Panel ARDL/NARDL** (provided by `pnardl`, `xtpqardl`, `xtcspqardl`)

The `nardl` package on CRAN implements basic nonlinear ARDL but lacks bootstrap
inference, Fourier terms, and panel extensions. The `dynamac` package provides
dynamic simulation from ARDL models but not bounds testing or nonlinear variants.

The `ardlverse` ecosystem thus represents the most comprehensive ARDL toolkit
available in R, covering linear, nonlinear, quantile, Fourier, bootstrap, and
panel extensions in a unified framework.

# Summary

The `ardlverse` ecosystem provides nine packages that collectively offer the most
comprehensive ARDL modeling toolkit in any statistical software environment. The
packages cover augmented ARDL, rolling/recursive ARDL, Fourier NARDL, bootstrap
NARDL, multiple-threshold NARDL, quantile ARDL, Fourier quantile ARDL, panel
NARDL, panel quantile ARDL, and cross-sectionally augmented panel quantile ARDL.
All packages are available on CRAN under GPL-3.

\bibliography{paper}
